Efficiently Handling Guards in the Low Level Design of RPython's Tracing JIT

### 2012-08-10 talk/vmil2012/paper.tex
a note about luajit

Mike Pall, the author of LuaJIT describes in a post to the lua-users mailing
list different technologies and techniques used in the implementation of
LuaJIT~\cite{Pall:2009}. Pall explains that guards in LuaJIT use a datastucture
called snapshots, similar to RPython's resume data, to store the information
about how to rebuild the state from a side-exit using the information in the
snapshot and the machine execution state. According to Pall~\cite{Pall:2009}
snapshots for guards in LuaJIT are associated with a large memory footprint.
The solution used in there is to store sparse snapshots, avoiding the creation
of snapshots for every guard to reduce memory pressure. Snapshots are only
created for guards after updates to the global state, after control flow points
from the original program and for guards that are likely to fail. As an outlook
Pall mentions the plans to switch to compressed snapshots to further reduce
redundancy.

@@

\section{Conclusion}
\label{sec:Conclusion}

\todo{conclusion}

>>>>

Mike Pall, the author of LuaJIT describes in a post to the lua-users mailing
list different technologies and techniques used in the implementation of
LuaJIT~\cite{Pall:2009}. Pall explains that guards in LuaJIT use a datastucture
called snapshots, similar to RPython's resume data, to store the information
about how to rebuild the state from a side-exit using the information in the
snapshot and the machine execution state. According to Pall~\cite{Pall:2009}
snapshots for guards in LuaJIT are associated with a large memory footprint.
The solution used in there is to store sparse snapshots, avoiding the creation
of snapshots for every guard to reduce memory pressure. Snapshots are only
created for guards after updates to the global state, after control flow points
from the original program and for guards that are likely to fail. As an outlook
Pall mentions the plans to switch to compressed snapshots to further reduce
redundancy. The approach of not creating snapshots at all for every guard is
orthogonal to the resume data compression presented in this paper and could be
reused within RPython to improve the memory usage further.

@@

\section{Conclusion}
\label{sec:Conclusion}
In this paper we have concentrated on guards, an operation typically found in
tracing just-in-time compilers and used to denote points of possible control
flow bifurcation in recorded traces.
We described how, based on the observation that guards are a frequent operation
in traces and that they do not fail often, guards have been implemented in the
high and low level components of RPython's tracing JIT compiler.

Finally we have presented experimental data collected using the standard PyPy
benchmark set to evaluate previous observations and assumptions. Our
experiments showed that, as previously assumed, guards are a very common
operation in traces. At the same time guards are associated with a high
overhead, because for all compiled guards high- and low information needs to be
stored to restore the execution state in case of a bail-out. The measurements
showed that the compression techniques used in PyPy effectively reduce the
overhead of guards, while it still produces a significant overhead. The results
also showed that guard failure is a local event, meaning that there are few
guards that fail at all, and ever fewer that fail very often, validating the
concept of reducing the overhead of successful guard checks while incurring an
additional overhead, besides leaving the optimized code, in the case of a
bailout having to decode the compressed state representation, which on the
other hand reduces the memory footprint storing data that is seldom used in a
compact manner.

\subsection{Future Work}
% subsection Future Work (end)
An aspect that requires further investigation is to experimentally validate
different of the trade-off decisions that went into the design of guard in
RPython's tracing JIT. Based on the observation that most guards do not fail
very often or at all while there are many guard and a few fail very often it
would be worth exploring if a more aggressive compression scheme for guards
would be worth the memory saving in contrast to the increased decoding
overhead. Based on the same observation we would like to explore the concept of
LuaJIT's sparse snapshots and its applicability to PyPy. Taking the idea of
data-driven optimization to investigate possible areas of optimization as was
done here in retrospective to validate a design seems a promising approach in
the context of tracing JITs.
